# Introduction

There are growing number of open source noncompartmental analysis R packages and we will compare, review and discuss them.

## In this chapter

We will see what `ncar` [@ncar2017] is and how it works.

## What is `ncar` R package?

Description: Conduct a noncompartmental analysis as closely as possible to the most widely used commercial software for pharmacokinetic analysis, i.e. 'Phoenix(R) WinNonlin(R)' <https://www.certara.com/software/pkpd-modeling-and-simulation/phoenix-winnonlin/>.

Some features include:
    
1. Use CDISC SDTM PP domain terms.
2. Automatic slope selection with the same criterion of WinNonlin(R)
3. Support both 'linear-up linear-down' and 'linear-up log-down' method
4. Calculate partial(interval) AUC with 'linear' or 'log' interpolation method
5. Perform a noncompartmental analysis of CDISC standardized pharmacokinetic dataset (.XPT)

For more details on noncompartmental analysis, see the reference: Gabrielsson J, Weiner D. Pharmacokinetic and Pharmacodynamic Data Analysis - Concepts and Applications. 5th ed. 2016. (ISBN:9198299107)

## How does it work?

```{r}
library(ncar)

# Theoph and Indometh data: dose in mg, conc in mg/L, time in h
NCA(Theoph, "Subject", "Time", "conc", Dose=320)
NCA(Indometh, "Subject", "time", "conc", Dose=25, AdmMode="Bolus")

iAUC = data.frame(Name=c("AUC[0-12h]","AUC[0-24h]"), Start=c(0,0), End=c(12,24)) ; iAUC
NCA(Theoph, "Subject", "Time", "conc", Dose=320, iAUC=iAUC)
NCA(Indometh, "Subject", "time", "conc", Dose=25, AdmMode="Bolus", iAUC=iAUC)
```

```{r}
NCA(Theoph, "Subject", "Time", "conc", Dose=320, Report="Text")
```

# Comparison of Noncompartmental analysis software

## In this chapter

We will briefly compare NCA softwares.

## Certara Phoenix WinNonLin

<https://www.certara.com/software/pkpd-modeling-and-simulation/phoenix-winnonlin/>

### Pros

- Validated for several years
- Industry standard
- Versatile unit setting
- Easy using by GUI 
- Generating plots supported

### Cons

- Expansive (~several thousand dollars)
- Not suitable for reproducible research
- CDISC SDTM not compatible (input and output)

## R package: PKNCA

<https://github.com/billdenney/pknca>

### Pros

- Open source and free of charge
- CDISC SDTM semi compatible (output)
- Calculate partial(interval) AUC with 'linear' or 'log' interpolation method but in a cumbersome way

### Cons
 
- CDISC SDTM not compatible (input)
- More tests required
- Experience with R language required
- Generating plots not supported for now (To be supported soon)


## R package: ncar

Extension of NonCompart for CDISC input

<https://github.com/asancpt/ncar>

### Pros

- Open source and free of charge
- CDISC SDTM fully compatible (input - DM, EX, PC)
- Output of CDISC SDTM PP domain terms 
- Calculate partial(interval) AUC with 'linear' or 'log' interpolation method

### Cons

- More tests required
- Experience with R language required
- Generating plots not supported for now (To be supported soon)

## R package: ncappc

NCA Calculation and Population PK Model Diagnosis [@Acharya201683]

<https://cran.r-project.org/web/packages/ncappc/index.html>
<https://www.ncbi.nlm.nih.gov/pubmed/27000291>



```{r}
#install.packages("ncappc")
library(ncappc)
```


## Kinetica

## PK module in R

## Scientist

## PKSolver

## R package: PK

Basic Non-Compartmental Pharmacokinetics

<https://cran.r-project.org/web/packages/PK/index.html>


```{r}
#install.packages("PK")
library(PK)
```

## Summary

```{r echo = FALSE}
knitr::kable(read.csv("comparison.csv", check.names = FALSE))
```

# Details


## In this chapter

We will give you details of each software and perform NCA with a dataset `Theoph`. In the end, we will compare C~max~ and AUC~inf~ calculated by each software or package.

## ncar

```{r}
# install.packages("ncar")
library(ncar)
NCA(Theoph, "Subject", "Time", "conc", Dose=320, Method = "Log") %>% select(CMAX, AUCIFO)
```

## PKNCA

```{r}
library(PKNCA)

my.conc <- PKNCAconc(as.data.frame(Theoph), conc~Time|Subject)
d.dose <- unique(datasets::Theoph[datasets::Theoph$Time == 0,
                                  c("Dose", "Time", "Subject")])
my.dose <- PKNCAdose(d.dose, Dose~Time|Subject)
my.data.automatic <- PKNCAdata(my.conc, my.dose)
my.results.automatic <- pk.nca(my.data.automatic)
my.results.automatic$result %>% filter(grepl(pattern = "cmax|aucinf", PPTESTCD)) %>% 
    arrange(PPTESTCD)
summary(my.results.automatic)
```

## ncappc

```{r eval = FALSE}
library(ncappc)

write.csv(Theoph %>% rename(ID = Subject, TIME = Time, DV = conc), 
          "Theoph.csv", row.names = FALSE)
ncappc(obsFile="Theoph.csv", psnOut = FALSE, noPlot = TRUE, printOut = TRUE, 
       method = "linear-log", evid = FALSE)
Theoph_ncappc <- read.delim("ncaOutput-.tsv", sep = "\t", check.names = FALSE)
Theoph_ncappc[ , c("Cmax (M.L^-3)", "AUCINF_obs (T*M.L^-3)")]
```

```{r echo = FALSE}
Theoph_ncappc <- read.delim("ncaOutput-.tsv", sep = "\t", check.names = FALSE)
Theoph_ncappc[ , c("Cmax (M.L^-3)", "AUCINF_obs (T*M.L^-3)")]
```


## Summary

`ncar` is good.
